---
geometry: margin = .9in
output: pdf_document
---

# EDA in R -- Stat 250, Winter 2022

```{r setup1, message=FALSE, include=FALSE}
# Don't change this part
knitr::opts_chunk$set(collapse=TRUE, comment=NULL)
```

## Loading R packages

```{r setup, message=FALSE}
# Load packages here:
library(mosaic)
library(tidyverse)
```

### Loading data from the web

```{r}
# Load the data here
bechdel <- read.csv("https://aloy.rbind.io/data/bechdel.csv")
```

#### Checkpoint:

How many cases are there in this data set?  

How many variables?  For each variable, identify its data type (e.g., categorical, quantitative).

```{r}
# Take a look at the data set
```



## Graphical exploration

### Create a bar chart of the Bechdel test results


```{r}
# Code for a bar chart here
```

Now, add a title and axis labels
```{r}
# Polish your bar chart here
```


####  Checkpoint:
Create a bar chart of the `binary` Bechdel test results.

```{r}
# Create a bar chart for binary
```


### Reorder a bar chart

```{r}
# Reorder the bars
```


### Create a histogram of domestic return on investment

```{r}
# Create a histogram here
```

### Change the number of bins in a histogram

```{r}
# Use 100 bins in the histogram
```

### Change the binwidth in a histogram

```{r}
# Use a binwidth of 10
```

#### Checkpoint: Add informative labels

```{r}
# Add a title and change the x-axis label
```

### Create a boxplot of domestic return on investment

```{r}
# Create a boxplot
```

```{r}
# Remove the y-axis entirely
```

### Transform a variable


```{r}
# Create log-transformed histograms and boxplots
```

### Side-by-side boxplots

```{r}
# Create side-by-side boxplots
```


### Scatterplots

```{r}
# Create a scatterplot here
```



## Numerical exploration


### Summary statistics for quantitative variables

Calculate summary stats for the budget in 2013 USD. What happens if you apply `summary` to the entire data frame?
```{r}
# Put your code here
```

Calculate the mean, median, IQR, standard deviation for `budget_2013`
```{r}
# Put your code here
```

Calculate the mean for `roi_dom`
```{r}
# Put your code here
```

Calculate the same summary statistics for `budget_2013`, but using the mosaic package syntax:

```{r}
# Put your code here
```

Now, use `favstats()`:
```{r}
# Put your code here
```

### Grouped summaries

Calculate a summary of `budget_2013` for each level of `clean_test` using the `favstats()` command:

```{r}
# Put your code here
```

Which category has the highest median budget? Which genre has the smallest first quartile (Q1) for budget?



### Summary statistics for categorical variables

Calculate the number of movies that passed of failed the Bechdel test using `tally()`:

```{r}
# Put your code here
```

Now look at the relative frequency distribution (proportions):

```{r}
# Put your code here
```


### Contingency tables

Break down the `binary` Bechdel test results by decade:

```{r}
# Put your code here
```


Breaking down `binary` by `decade_code`, calculating the proportion of movies that pass the Bechdel test by decade:

```{r}
# Put your code here
```

Add margins:

```{r}
# Put your code here
```



## Your turn

#### Calculate the number of movies released in each year.


```{r yt1}
# insert your code here
```

(Add your description and/or answer in plain text here.)


#### Describe the distribution of `budget_2013` in terms of its center, shape, and  spread, including any plots you use.

```{r yt2}
# insert your code here
```

(Add your description and/or answer in plain text here.)



#### Using numerical summaries and a side-by-side boxplot, determine if movie budgets have changed over the decades.

```{r yt3}
# insert your code here
```

(Add your description and/or answer in plain text here.)



#### Do movies that pass the Bechdel test tend to have higher or lower budgets than movies that fail the Bechdeul test? 

```{r yt4}
# insert your code here
```

(Add your description and/or answer in plain text here.)



